### 2025-05-07 thesis/thesis.tex
Write about pre trained language models

% 1/3 Vorarbeit, Literature Review
% 1/3 Decomp und System Design
% 1/3 Implementation und Evaluation

@@

\subsection{Limitations of Neural Approaches}

@@

% PLM approaches
% RESDSQL, Graphix-T5
% T5 fine-tuning etc.

% https://arxiv.org/abs/2109.05093

>>>>

% 1/3 Vorarbeit, Literature Review
% 1/3 Theory, Decomp und System Design
% 1/3 Implementation und Evaluation

@@

\subsubsection{Limitations of Neural Approaches}

@@

The advantages of combining specialized neural networks with general-purpose pre-trained language models
led to a pivotal point in the NL2SQL research field towards focusing increasingly on the application of pre-trained
language models for NLIDBs. Models like BERT or T5 offer noticable performance improvements (especially when it
comes to language understanding) over specialized NL2SQL networks due to training happening on unrestrained amounts
of natural language data, instead of pure NL2SQL datasets which are often fairly limited in size and therfore natural
language use — \textsc{Spider2.0} which is a contemprary NL2SQL benchmark consists of just 632 real-world questions 
\citep{SPIDER2}. Thus PLM-based (or at least augumented) NL2SQL systems can observe dramatic performance improvements
through the language models's ability to understand patterns and identify semantic relationships of natural language
query elements.

\subsubsection{Early Pre-trained Language Model Adaptations}

The above outlined benefits have led to concrete research efforts focusing on the question whether the sole application
of pre-trained language models could outperform neural state-of-the-art approaches — which often implicitly require
a far more sophisticate architecture when it comes to natural language analysis. The application of T5-3B for NL2SQL
tasks which yielded 71.4\% execution accuracy presented a breakthrough in this research domain \citep{T2SQL-LLM-Bench-3}.
This demonstrated that PLMs are correctly positioned to significantly outperform existing approaches.

In the time of emerging PLM applicationm \textsc{Grappa} was introduced by \citeauthor*{GRAPPA} in \citeyear{GRAPPA} —
a novel grammar-augmented pre-training approach built on RoBERTa\textsubscript{\tiny{LARGE}} (a derivative model from BERT).
It generates synthetic training data (ie. natural language and sql pairs) using a synchronous context-free grammar (SCFG)
which analyses and identifies patterns in natural language queries that can be used as templates for sythesizing training
data. The specialized pre-training helps \textsc{Grappa} to establish a robust connection between natural-language and
database schema elements, showing significant improvements on existing approaches on multiple contemporary benchmarks like 
\textsc{Spider} and \textsc{WikiSQL} \citep{GRAPPA}.

Several NL2SQL approaches in this era focused on \textit{schema understanding} and \textit{schema linking} — the 
generalizability of PLMs required advanced techniques on ensuring that models both understand the semantic intent
of users when querying and correctly identify database schema elements in natural language queries. Thus improving
semantic accuracy of generated SQL queries. \textsc{StruG} (Structure-Grounded-Pretraining) was introduced in
\citeyear{STRUG} by \citeauthor*{STRUG} and presented a novel pretraining approach that improves model abilities when it comes to
\textit{schema linking}, it separates the problem in three facets: column grounding, value grounding and
column-value mapping. In direct comparison with \textsc{Grappa}, \textsc{StruG} achieves similar performance
while being significantly cheaper to train \citep{STRUG}.

\subsubsection{Intermediate Pre-trained Language Model Approaches}

% RYANSQL, GAZP
% RESDSQL, GRAPHIX, NATSQL

% T5 fine-tuning etc.
% https://arxiv.org/pdf/2302.05965

\subsubsection{Constrained Decoding and Ranking Techniques}

% PICARD etc.

\subsubsection{Comparison with Large Language Models}
